---
title: "Beer Production"
output: html_notebook
---

This is a R notebook for the TidyTuesday dataset on Beer Production. This notebook is following the work of David Robinson, as I gain expereince with R and working with Data Visualization. First, lets grab our datasets.
```{r}
library(tidyverse)
theme_set(theme_light())
brewing_materials <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv')
beer_taxed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_taxed.csv')
brewer_size <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewer_size.csv')
beer_states <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_states.csv')
```

To start with, what ingredients are used in beer? This is all the data presented for one month
```{r}
brewing_materials %>%
  filter(year == max(year), month == max(month)) %>%
  ggplot(aes(type, month_current, fill = material_type)) +
  geom_col() + 
  coord_flip()
```

What about looking at just the materials used in one month?
```{r}
brewing_materials %>%
  filter(year == max(year), month == max(month)) %>%
  filter(!str_detect(material_type, "Total")) %>%
  mutate(type = fct_reorder(type, month_current)) %>%
  ggplot(aes(type, month_current, fill = material_type)) +
  geom_col() +
  coord_flip()
```

Now lets look at the usage over time. We will change the date to be the first of the month.
```{r}
library(lubridate)

brewing_materials <- brewing_materials %>%
  mutate(date = ymd(paste(year, month, 1)))

brewing_materials %>%
  filter(!str_detect(material_type, "Total")) %>%
  ggplot(aes(date, month_current, fill = type)) + 
  geom_col()
```
It looks like Rice and Rice products do not exist in the data after 2015, lets remove these datapoints.

```{r}
brewing_materials <- brewing_materials %>%
  filter(year < 2016)

brewing_materials %>%
  filter(!str_detect(material_type, "Total")) %>%
  mutate(type = fct_reorder(type, month_current, sum)) %>%
  ggplot(aes(date, month_current, fill = type)) + 
  geom_col() + 
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Time", 
       y = "Pounds used in beer production",
       fill = "Material")
```

Now lets only look at the total products used
```{r}
brewing_materials %>%
  filter(str_detect(material_type, "Total.*products")) %>%
  mutate(type = fct_reorder(type, month_current, sum)) %>%
  ggplot(aes(date, month_current, fill = type)) + 
  geom_col() + 
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Time", 
       y = "Pounds used in beer production",
       fill = "Material")
```

```{r}
devtools::install_github("ramnathv/tidymetrics")
library(tidymetrics)
```

```{r}
brewing_materials %>%
  filter(!str_detect(material_type, "Total")) %>%
  cross_by_dimensions(material_type, type) %>%
  distinct(material_type, type)


brewing_materials %>%
  rename(material = type) %>%
  filter(!str_detect(material_type, "Total")) %>%
  cross_by_dimensions(material_type, material) %>%
  summarize(nb_pounds = sum(month_current)) %>%
  View()
```

brewing_summarized allows us to see the total pattern of beer production by month.
```{r}
brewing_summarized <- brewing_materials %>%
  rename(material = type) %>%
  filter(!str_detect(material_type, "Total")) %>%
  cross_by_dimensions(material_type, material) %>%
  cross_by_periods(c("month", "quarter", "year")) %>%
  summarize(nb_pounds = sum(month_current))

brewing_summarized %>%
  filter(material_type == "All", material == "All", period == "month") %>%
  ggplot(aes(date, nb_pounds)) + 
  geom_col()
  
```

... or by quarter
```{r}
brewing_summarized %>%
  filter(material_type == "All", material == "All", period == "quarter") %>%
  ggplot(aes(date, nb_pounds)) +
  geom_col()
```
We are keeping the same filtering, but instead we are looking at a different cross section of this data. The below graph is looking at the useage of grain products and non-grain products, by year.
```{r}
brewing_summarized %>%
  filter(material_type != "All", material == "All", period == "year") %>%
  ggplot(aes(date, nb_pounds, fill = material_type)) +
  geom_col()
```
If we wanted to look at the material instead,
```{r}
brewing_summarized %>%
  filter(material_type == "All", material != "All", period == "year") %>%
  ggplot(aes(date, nb_pounds, fill = material)) +
  geom_col()
```
We can also filter this by quarter
```{r}
brewing_summarized %>%
  filter(material_type == "All", material != "All", period == "quarter") %>%
  ggplot(aes(date, nb_pounds, fill = material)) + 
  geom_col()
```


