---
title: "Beer Production"
output: html_notebook
---

This is a R notebook for the TidyTuesday dataset on Beer Production. This notebook is following the work of David Robinson, as I gain expereince with R and working with Data Visualization. First, lets grab our datasets.
```{r}
library(tidyverse)
theme_set(theme_light())
brewing_materials <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv')
beer_taxed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_taxed.csv')
brewer_size <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewer_size.csv')
beer_states <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_states.csv')
```

To start with, what ingredients are used in beer? This is all the data presented for one month
```{r}
brewing_materials %>%
  filter(year == max(year), month == max(month)) %>%
  ggplot(aes(type, month_current, fill = material_type)) +
  geom_col() + 
  coord_flip()
```

What about looking at just the materials used in one month?
```{r}
brewing_materials %>%
  filter(year == max(year), month == max(month)) %>%
  filter(!str_detect(material_type, "Total")) %>%
  mutate(type = fct_reorder(type, month_current)) %>%
  ggplot(aes(type, month_current, fill = material_type)) +
  geom_col() +
  coord_flip()
```

Now lets look at the usage over time. We will change the date to be the first of the month.
```{r}
library(lubridate)

brewing_materials <- brewing_materials %>%
  mutate(date = ymd(paste(year, month, 1)))

brewing_materials %>%
  filter(!str_detect(material_type, "Total")) %>%
  ggplot(aes(date, month_current, fill = type)) + 
  geom_col()
```
It looks like Rice and Rice products do not exist in the data after 2015, lets remove these datapoints.

```{r}
brewing_materials <- brewing_materials %>%
  filter(year < 2016)

brewing_materials %>%
  filter(!str_detect(material_type, "Total")) %>%
  mutate(type = fct_reorder(type, month_current, sum)) %>%
  ggplot(aes(date, month_current, fill = type)) + 
  geom_col() + 
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Time", 
       y = "Pounds used in beer production",
       fill = "Material")
```

Now lets only look at the total products used
```{r}
brewing_materials %>%
  filter(str_detect(material_type, "Total.*products")) %>%
  mutate(type = fct_reorder(type, month_current, sum)) %>%
  ggplot(aes(date, month_current, fill = type)) + 
  geom_col() + 
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Time", 
       y = "Pounds used in beer production",
       fill = "Material")
```

```{r}
devtools::install_github("ramnathv/tidymetrics")
library(tidymetrics)
```

```{r}
brewing_materials %>%
  filter(!str_detect(material_type, "Total")) %>%
  cross_by_dimensions(material_type, type) %>%
  distinct(material_type, type)


brewing_materials %>%
  rename(material = type) %>%
  filter(!str_detect(material_type, "Total")) %>%
  cross_by_dimensions(material_type, material) %>%
  summarize(nb_pounds = sum(month_current)) %>%
  View()
```

brewing_summarized allows us to see the total pattern of beer production by month.
```{r}
brewing_summarized <- brewing_materials %>%
  rename(material = type) %>%
  filter(!str_detect(material_type, "Total")) %>%
  cross_by_dimensions(material_type, material) %>%
  cross_by_periods(c("month", "quarter", "year")) %>%
  summarize(nb_pounds = sum(month_current))

brewing_summarized %>%
  filter(material_type == "All", material == "All", period == "month") %>%
  ggplot(aes(date, nb_pounds)) + 
  geom_col()
  
```

... or by quarter
```{r}
brewing_summarized %>%
  filter(material_type == "All", material == "All", period == "quarter") %>%
  ggplot(aes(date, nb_pounds)) +
  geom_col()
```
We are keeping the same filtering, but instead we are looking at a different cross section of this data. The below graph is looking at the useage of grain products and non-grain products, by year.
```{r}
brewing_summarized %>%
  filter(material_type != "All", material == "All", period == "year") %>%
  ggplot(aes(date, nb_pounds, fill = material_type)) +
  geom_col()
```
If we wanted to look at the material instead,
```{r}
brewing_summarized %>%
  filter(material_type == "All", material != "All", period == "year") %>%
  ggplot(aes(date, nb_pounds, fill = material)) +
  geom_col()
```
We can also filter this by quarter
```{r}
brewing_summarized %>%
  filter(material_type == "All", material != "All", period == "quarter") %>%
  ggplot(aes(date, nb_pounds, fill = material)) + 
  geom_col()
```

Lets now look at the number of barrels that come from larger breweries vs smaller breweries.First we will look at the total number of barrels produced.
```{r}
brewer_size %>%
  ggplot(aes(year, total_barrels)) + 
  geom_col()
```

```{r}
brewer_size %>%
  filter(brewer_size != "Total") %>%
  ggplot(aes(year, total_barrels, fill = brewer_size)) + 
  geom_col()
```

Now lets fix the ordering of the legend
```{r}
brewer_size %>%
  filter(brewer_size != "Total", !is.na(total_barrels)) %>%
  mutate(brewer_size = fct_lump(brewer_size, 5, w = total_barrels),
         barrel_number = coalesce(parse_number(as.character(brewer_size)), 1),
         brewer_size = fct_reorder(brewer_size, barrel_number)) %>%
  ggplot(aes(year, total_barrels, fill = brewer_size)) + 
  geom_col()
```

Let's display the brewing materials graph in a different way. In this graph we can look at the seasonal changes in beer production.
```{r}
brewing_materials %>%
  filter(material_type == "Total Used") %>%
  ggplot(aes(month, month_current, color = factor(year))) + 
  geom_line() + 
  expand_limits(y = 0)
```
Where is beer produced?
```{r}
beer_states %>%
  count(type, sort = TRUE, wt = barrels)
```

Now let's make a map of beer production
```{r}
beer_states %>%
  group_by(year) %>%
  summarize(barrels = sum(barrels, na.rm = TRUE))
  
# Who consumes beer on premises?
beer_states %>%
  filter(type == "On Premises", year == max(year), state != "total") %>%
  arrange(desc(barrels))
```

Let's look at this as a percentage of beer each state makes for on premise consumption.
```{r}
beer_states %>%
  filter(year == max(year), state != "total") %>%
  group_by(state) %>%
  mutate(percent = barrels / sum(barrels)) %>%
  filter(type == "On Premises") %>%
  arrange(desc(percent))
```
What about for bottles and cans
```{r}
beer_states %>%
  filter(year == max(year), state != "total") %>%
  group_by(state) %>%
  mutate(percent = barrels / sum(barrels)) %>%
  filter(type == "Bottles and Cans") %>%
  arrange(desc(percent))
```
Now lets see what percentage of beer produced is for kegs and barrels.
```{r}
beer_states %>%
  filter(year == max(year), state != "total") %>%
  group_by(state) %>%
  mutate(percent = barrels / sum(barrels)) %>%
  filter(type == "Kegs and Barrels") %>%
  arrange(desc(percent))
```
```{r}
state_percents_2019 <- beer_states %>%
  filter(year == max(year), state != "total") %>%
  group_by(state) %>%
  mutate(percent = barrels / sum(barrels)) %>%
  ungroup()

state_percents_2019 %>%
  filter(type == "On Premises") %>%
  arrange(desc(percent))
```
```{r}
library(maps)
library(sf)

states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
```

```{r}
state_percents_2019 %>%
  filter(type == "On Premises") %>%
  mutate(ID = str_to_lower(state.name[match(state, state.abb)])) %>%
  inner_join(states, by = "ID") %>%
  ggplot(aes(geometry = geom, fill = percent)) + 
  geom_sf() + 
  scale_fill_gradient2(low = "blue", high = "red", midpoint = .5,
                       labels = scales::percent) +
  ggthemes::theme_map() + 
  labs(title = "Percentage of produced beer consumed on premise, in 2019",
       fill = "% consumed on prem.")
```

```{r}
states_joined <- state_percents_2019 %>%
  mutate(ID = str_to_lower(state.name[match(state, state.abb)])) %>%
  inner_join(states, by = "ID")

states_joined %>%
  ggplot(aes(geometry = geom, fill = percent)) + 
  geom_sf() +
  scale_fill_gradient2(low = "blue", high = "red", midpoint = 0.5,
                       labels = scales::percent) + 
  facet_wrap(~ type, nrow = 2) +
  ggthemes::theme_map() + 
  theme(legend.position = "bottom") + 
  labs(title = "How each state consumed their beer in  2019", 
       fill = "% consumed")
  
```

Animated graph?
```{r}
library(gganimate)

beer_states %>%
  filter(state != "total") %>%
  group_by(state, year) %>%
  mutate(percent = barrels / sum(barrels)) %>%
  ungroup() %>%
  filter(type == "On Premises") %>%
  mutate(ID = str_to_lower(state.name[match(state, state.abb)])) %>%
  inner_join(states, by = "ID") %>%
  ggplot(aes(geometry = geom, fill = percent)) +
  geom_sf() +
  transition_manual(year) +
  # facet_wrap(~ type, nrow = 2) +
  scale_fill_gradient2(low = "blue", high = "red", midpoint = 0.5,
                       labels = scales::percent) +
  ggthemes::theme_map() +
  theme(legend.position = "bottom") + 
  labs(title = "How is beer consumed within each state",
       fill = "% consumed") 
```

